Monitoring forest health using hyperspectral imagery: Does feature selection improve the performance of machine-learning techniques?

### 2021-06-06 docs/00-manuscripts/ieee/ieee-manuscript.tex
slighlty rephrase PCA correlation analysis

% \documentclass[letterpaper, onecolumn, peerreview]{IEEEtran}
\documentclass[journal]{IEEEtran}

@@

% *** CITATION PACKAGES ***
%
\usepackage{cite}
\usepackage[hidelinks]{hyperref}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
%\cite{} output to follow that of the IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% 'compressed/ranged'. e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
%\cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/pkg/cite
% The documentation is contained in the cite.sty file itself.

@@

In total the benchmarking grid consisted of 156 experiments (6 feature sets $\times$ 3 ML algorithms $\times$ 8 feature-selection methods and for the L1/L2 models, 6 feature sets $\times$ 2 models.
The selected hyperparameter settings are shown in \autoref{tab:hyperparameter_limits}.
In addition the code which generated the matrix is available in the research compendium of this study (\url{10.5281/zenodo.2635403}).

@@

The idea of splitting the features into different sets originated from the question whether feature-engineered indices derived from reflectance values have a positive effect on model performance.
\cite{pena2017} is an exemplary study which used this approach in a spectro-temporal setting.
Benchmarking learners on these feature sets while keeping all other variables such as model type, tuning strategy and partitioning method constant makes it possible to draw conclusions on their individual impact.
However, rather than only looking at these three groups also combinations of such were taken into account:

@@

The PCA-based correlation analysis showed that, depending on the feature set, between two (HR) and 42 (NRI) \ac{PC} are needed to explain 95\% of variance.
The fewer PCs are needed to reach a high proportion of variance explained the more similar the individual features are to each other.
Hence, a high correlation can be observed for feature set HR which reaches 96\% with only two PCs.
Feature sets NRI and VI are more diverse and require 42 (NRI) and 12 (VI) PCs to explain 95\% of the variance in the data, respectively.

>>>>

% \documentclass[letterpaper, onecolumn, peerreview]{IEEEtran}
\documentclass[final]{IEEEtran}

@@

% *** CITATION PACKAGES ***
%
\usepackage[noadjust]{cite}
\usepackage[hidelinks]{hyperref}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
%\cite{} output to follow that of the IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% 'compressed/ranged'. e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
%\cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/pkg/cite
% The documentation is contained in the cite.sty file itself.

@@

In total the benchmarking grid consisted of 156 experiments (6 feature sets $\times$ 3 ML algorithms $\times$ 8 feature-selection methods and for the L1/L2 models, 6 feature sets $\times$ 2 models.
The selected hyperparameter settings are shown in appendix \autoref{tab:hyperparameter_limits}.
In addition the code which generated the matrix is available in the research compendium of this study (\url{10.5281/zenodo.2635403}).

@@

The idea of splitting the features into different sets originated from the question whether feature-engineered indices derived from reflectance values have a positive effect on model performance.
Peña et al. 2017 \cite{pena2017} is an exemplary study which used this approach in a spectro-temporal setting.
Benchmarking learners on these feature sets while keeping all other variables such as model type, tuning strategy and partitioning method constant makes it possible to draw conclusions on their individual impact.
However, rather than only looking at these three groups also combinations of such were taken into account:

@@

PCA was used for a correlation analysis of the feature sets by inspecting the accumulated percentages of the principal components.
The fewer PCs are needed to reach a high proportion of variance explained the more similar the individual features are to each other.
Results showed that, depending on the feature set, between two (HR) and 42 (NRI) \ac{PC} are needed to explain 95\% of variance.
Hence, a high correlation can be observed for feature set HR which reaches 96\% with only two PCs.
Feature sets NRI and VI are more diverse and require 42 (NRI) and 12 (VI) PCs to explain 95\% of the variance in the data, respectively.
